Bringing the Power of eBPF to Open vSwitch

### 2018-10-21 paper.tex
add afxdp txrx

All the packet data buffers used in TX/RX ring are allocated from a specific
memory area called UMEM which consists of a number of fixed size chunks.
A descriptor in RX/TX ring points to the element in UMEM by its address.
The address is simply an offset within the entire UMEM memory region.
The UMEM also has two rings: the FILL ring and the COMPLETION ring.
The Fill ring is used by the application to send down addr for the kernel
to fill in with RX packet data. References to these frames will then
appear in the RX ring once each packet has been received. The 
completion ring, on the other hand, contains frame addr that the 
kernel has transmitted completely and can now be used again by user
space, for either TX or RX.

@@

@@

\subsection{AF\_XDP netdev configuration}
ovs-vswitchd controls and manages the XDP program and maps.
When users attach a AF\_XDP netdev to the OVS bridge, the netdev-afxdp
does the following steps:
We implemnet most of the logic in lib/netdev-linux.c
{
\begin{verbatim}
    ovs-vsctl add-br br0
    ovs-vsctl add-port br0 eth0 -- \
        set int eth0 type="afxdp"
\end{verbatim}
}
\begin{enumerate}
\item Attach a XDP program to the netdev queue:
The XDP program OVS attaches is fixed and consists
of only a few line of code, which receives the packet and
calls the \texttt{bpf\_redirect\_map()} helper function with
the XSK eBPF maps and key equals zero.

\item Load and configure the XSK eBPF map:
The XSK eBPF maps consists of key value pairs, where key is an u32 index and value is
the netdev's queue ID.  In our case, we simply program one entry to the map with
key equals zero and the file descriptor, fd, of the XSK as its value.
As a result, the XDP program calling \texttt{bpf\_redirect\_map} with key=0 will
forwards the packet to the XSK.
\end{enumerate}

\subsection{Packet Reception and Transmission}

When receiving the packets from XSK, netdev-afxdp implements the following
operations:
\begin{enumeration}
\item Get N free elements from UMEM
\item Fille

\end{enumeration}
When sending out the packets

\subsection{Performance Evaluation}
PMD
HugePage
UMEM element list
Batching

>>>>

% concept
Application still HW agnostic with ZC
Each application gets its own packet buffer and TX/RX descriptor rings
Packet buffers can be shared if desired
  TX/RX descriptor rings always private to process
% AF\_XDP has 3 mode
% zero copy, native mode,
XDP\_SKB:
– Works on any netdevice using sockets and generic XDP path
XDP\_DRV:
– Works on any device with XDP support (all three NDOs)
XDP\_DRV + ZC:
– Need buffer allocator support in driver + a new NDO for TX
All the packet data buffers used in TX/RX ring are allocated from a specific
memory area called UMEM which consists of a number of fixed size chunks.
A descriptor in RX/TX ring points to the element in UMEM by its address.
The address is simply an offset within the entire UMEM memory region.
The UMEM also has two rings: the FILL ring and the COMPLETION ring.
The Fill ring is used by the application to send down addr for the kernel
to fill in with RX packet data. References to these frames will then
appear in the RX ring once each packet has been received. The 
completion ring, on the other hand, contains frame addr that the 
kernel has transmitted completely and can now be used again by user
space, for either TX or RX.

@@

\subsection{AF\_XDP netdev configuration}
ovs-vswitchd controls and manages the XDP program and maps.
When users attach a AF\_XDP netdev to the OVS bridge, the netdev-afxdp
does the following steps:
We implemnet most of the logic in lib/netdev-linux.c
{
\begin{verbatim}
    ovs-vsctl add-br br0
    ovs-vsctl add-port br0 eth0 -- \
        set int eth0 type="afxdp"
\end{verbatim}
}

\begin{enumerate}
\item Attach a XDP program to the netdev queue:
The XDP program OVS attaches is fixed and consists
of only a few line of code, which receives the packet and
calls the \texttt{bpf\_redirect\_map()} helper function with
the XSK eBPF maps and key equals zero.

\item Load and configure the XSK eBPF map:
The XSK eBPF maps consists of key value pairs, where key is an u32 index and value is
the netdev's queue ID.  In our case, we simply program one entry to the map with
key equals zero and the file descriptor, fd, of the XSK as its value.
As a result, the XDP program calling \texttt{bpf\_redirect\_map} with key=0 will
forwards the packet to the XSK.

\item Initialize the FILL queue
Before any packet arrival, the rest three queues: RX, TX and COMPLETION
remains the same
\end{enumerate}

memory allocation:

% 1 UMEM per XSK per queue, no sharing
For each netdev, ovs-vswitchd attach XSK to each 
For each XSK
memorypool
We implement a UMEM freelist to keep track of the used/unused element

To handle multiple 
\subsection{Packet Reception and Transmission}
\begin{figure}
{\scriptsize
\begin{verbatim}
UMEM: total 8 elements, each has 2K chunk size
                 +-------------------------------+
                 |   |   |   |   |   |   |   |   |
                 +-------------------------------+
            addr:  1   2   3   4   5   6   7  8 
(1) Initial Stat
           -------------------          UMEM freelist
FILL Qu  ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
           -------------------
           -------------------
RX ring  ...|   |   |   |   | ...
           -------------------
(2) 4 Packets Arrive
           -------------------          UMEM freelist
FILL Qu  ...|   |   |   |   | ...       {5, 6, 7, 8}
          --------------------
           -------------------
RX ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(3) Get 4 free UMEM elemet and refill
           -------------------          UMEM freelist
FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {} 
           -------------------
           -------------------
RX ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(4) Pass packets {1, 2, 3, 4} to dpif-netdev
    for parse, lookup, and actions.

(5) Recycle to UMEM freelist, goto step (2)
            +-------------------        UMEM freelist
FILL Qu  ...| 5 | 6 | 7 | 8 | ...       {1, 2, 3, 4}
            +------------------
           -------------------
RX ring  ...|   |   |   |   | ...
           -------------------
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of how OVS programs the FILL queue and RX ring when
processing incoming AF\_XDP packets.}
\label{afxdprx}
\vspace{-1.0em}
\end{figure}
Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the FILL queue and
the RX ring for receving packets from XSK.  For simplicity to demonstrate
how AF\_XDP works, assume that there are only eight UMEM buffers, with each buffer's
size equals 2KB.  Initially at step 1, ovs-vswitchd fills four empty UMEM elem into
the FILL queue and waits for incoming packets.  When there are incoming packets,
at step 2, the FILL queue's four buffer elems have been consumed and moved
to the RX ring.  In order to keep receiving packets, ovs-vswitchd gets another
four empty UMEM elems from thte UMEM freelist, and fills into FILL queue (step 3).
Then ovs-vswitchd creates the metadata needed for the four packet
buffer, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
the dpif-netdev layer for parse, lookup and action executions.
Finally, when ovs-vswitchd finishes processing the UMEM buffer, a recycle
mechanism is triggered to place these buffer back to UMEM freelist (step 5).
Step 5 makes sure that there are always available elems in FILL queue, so
that the underlying XDP program in kernel can keep placing packets into
the FILL queue while the userspace ovs-vswitchd is processing the previous
received packets.
When step 5 finishes, ovs-vswitchd goes back to step 2, waiting and processing
new packets.

\begin{figure}
{\scriptsize
\begin{verbatim}
 dpif-netdev has 4 packets at UMEM:{1, 2, 3, 4}
 from eth1 and plan to send to eth2.
(1) Get 4 free element from eth2 UMEM, copy packet data to it
(2) Create TX descriptors and place into TX ring
           -------------------        eth2 UMEM freelist
TX ring  ...| 4 | 5 | 6 | 7 | ...       {1, 2, 3, 8}
           -------------------
           -------------------
COMPL q  ...|   |   |   |   | ...
           -------------------
(3) Issue send syscall to eth2's XSK
(4) Packet transmission is completed
           -------------------        eth2 UMEM freelist
TX ring  ...|   |   |   |   | ...       {1, 2, 3, 8}
           -------------------
           -------------------
COMPL q  ...| 4 | 5 | 6 | 7 | ...
           -------------------
(5) Recycle to eth2 UMEM
           -------------------        eth2 UMEM freelist
TX ring  ...|   |   |   |   | ...       {1, 2, 3, 4,
           -------------------           5, 6, 7, 8}
           -------------------
COMPL q  ...|   |   |   |   | ...
           -------------------
(6) Recycle {1, 2, 3, 4} to eth1's UMEM 
\end{verbatim}
}
\vspace{-1.0em}
\caption{An example of how OVS programs the COMPLETION queue and TX ring when
processing incoming AF\_XDP packets from one netdev and sending to another netdev.}
\label{afxdptx}
\vspace{-1.0em}
\end{figure}

Figure~\ref{afxdptx} shows the process for sending packets to the XSK.
Assuming that ovs-vswitch has a a bridge and two ports: eth1 and eth2.
And there is a flow entry showing
\texttt{in\_port=eth1, ..., actions=..., output:eth2}, meaning to forwarding
packets received from eth1 to eth2.
Assuming that both eth1 and eth2 are configured with AF\_XDP support.

Initially, assume ovs-vswitchd receives four packets {1, 2, 3, 4} from eth1's XSK
To send to eth2 using its XSK, ovs-vswitchd first gets four packet buffers,
{4, 5, 6, 7}, from eth2's UMEM freelist and copy the packet data to their
packet buffer in UMEM.  Then TX descriptors for the four packets are
created and placed into eth2's TX ring (step 2).
At step 3, send syscall is issued to signal the kernel to start transmit.
As the send in XSK is asynchronos and send syscall only return return zero when
no error, ovs-vswitch has to poll the COMPLETION queue to make sure these four
packets have been sent out (step 4).
Once the four packets' descriptors shown up at the COMPLETION ring, at step 5,
ovs-vswitchd recycles their UMEM elememnts back to the eth2's UMEM freelist.
In addition, the original four packet buffer from eth1, {1, 2, 3, 4}, is
recycled back to eth1's UMEM freelist.

\subsection{Performance Evaluation}
When sending and receivng packets using AF\_XDP with OVS, two CPUs show
100\% utilitization:
\begin{itemize}
\item ksoftirqd:
\item ovs-vswitchd:
\end{itemize}
PMD
HugePage
UMEM element list
Batching
2-CPU operation
non-blocking send
non-block receive

@@
